Measuring Individual Partisan Harm in Redistricting

### 2021-09-11 paper/paper.Rmd
merge

---
title: Measuring Individual Partisan Harm in Redistricting
authors:
  - name: Cory McCartan
    thanks: Footnote stuff
    department: Department of Statistics
    affiliation: Harvard University
    location: Cambridge, MA
    email: cmccartan@fas.harvard.edu
  - name: Christopher T. Kenny
    department: Department of Government
    affiliation: Harvard University
    location: Cambridge, MA
    email: christopherkenny@fas.harvard.edu
abstract: |
  Enter the text of your abstract here.
keywords:
  - partisan fairness
  - representation
  - redistricting simulation
  - individual standing
citation_package: biblatex
bibliography: references.bib
biblio-style: apa
output: 
  rticles::arxiv_article:
    keep_tex: true
    includes:
      in_header: "header.tex"
    latex_engine: pdflatex
---

@@

Decades of literature on redistricting have focused on the fairness to parties.
This literature defers to symmetric outcomes for parties, commonly referred to as partisan symmetry \cite{a:bunch:of:king}.
Partisan symmetry is a useful idea which greatly improves academic understanding of the redistricting process.
We Why is it hard?
Understanding the redistricting process requires a feasible and reliable understanding of the counterfactual set of redistricting plans.
Early methodologies for simulating plans relied on randomness to generate many different plans and enhanced understanding of redistricting greatly \\citep{ciri:etal:00, chen:rodd:13, defO:duch:solo:21.
These methods greatly enhanced the understanding of redistricting within contexts of a particular state with set populations.
Yet, without a known target distribution, these heuristic approaches to simulation fall do not offer necessary guarantees for confidence in the simulated plans to represent the total space of possible plans \cite{fife:etal:20b}.
Recent advances in Markov Chain methods as applied to redistricting allow us move beyond simulation towards sampling plans from this definable space \\cite{fife:etal:20, mcca:imai:20, autr:etal:21}.

Using sampling methods, we can then develop a reliable counterfactual that can be tuned to consider the samping constraints imposed by the laws and practices of a state.
This counterfactual allows us to characterize the likely plans from a state.
Often, we can would then call this an ensemble analysis, which is a flexible approach to sampling redistricting plans by simulation, relying on large numbers of plans for stability \cite{matt:etal:21} or sufficient diversity of nearly-independent plans \cite{kenn:etal:forth}.
using this counterfactual, we thus create measures of harm, utility, and fairness which are based on a single plan compared to a set of other reasonable and feasible plans which encorporate the necessary legal constraints to make a plan viable.

We offer a reminder that geographic context is key to understanding common problems in political science..
Elections themselves are geographically determined.
Recent research offers important resolution of the degree of partisan sorting in the US \cite{brow:enos:something}.
By combining those understandings with the necessities of redistricting, we are able to expand the localized aspect of harm.
Further, we suggest that while the efficiency gap may not incorporate geographic information, it often helps capture individual harm better than other common measures.
This is a great benefit as an indicator to detect extreme gerrymanders in cases where sufficiently local geographic data to simulate plans is not available.\
We further clarify that common existing measures of partisan fairness -- while they may capture the difficulties offered to a party -- do not necessarily capture the representational harms incurred by individuals who live in districts which do not at all represent their preferences.

>>>>

---
title: Measuring Individual Partisan Harm in Redistricting
authors:
  - name: Cory McCartan
    thanks: Footnote stuff
    department: Department of Statistics
    affiliation: Harvard University
    location: Cambridge, MA
    email: cmccartan@fas.harvard.edu
  - name: Christopher T. Kenny
    department: Department of Government
    affiliation: Harvard University
    location: Cambridge, MA
    email: christopherkenny@fas.harvard.edu
abstract: |
  Enter the text of your abstract here.
keywords:
  - partisan fairness
  - representation
  - redistricting simulation
  - individual standing
bibliography: references.bib
biblio-style: apalike
output: 
  rticles::arxiv_article:
    keep_tex: true
    includes:
      in_header: "header.tex"
    latex_engine: pdflatex
---

@@

Decades of literature on redistricting have focused on the fairness to parties.
This literature defers to symmetric outcomes for parties, commonly referred to as partisan symmetry [@katz2020].
Partisan symmetry is a useful idea which greatly improves academic understanding of the redistricting process.
Why is it hard?
Understanding the redistricting process requires a feasible and reliable understanding of the counterfactual set of redistricting plans.
Early methodologies for simulating plans relied on randomness to generate many different plans and enhanced understanding of redistricting greatly [@ciri:etal:00; @chen:rodd:13; @defO:duch:solo:21].
These methods greatly enhanced the understanding of redistricting within contexts of a particular state with set populations.
Yet, without a known target distribution, these heuristic approaches to simulation fall do not offer necessary guarantees for confidence in the simulated plans to represent the total space of possible plans [@fifield2020essential].
Recent advances in Markov Chain methods as applied to redistricting allow us move beyond simulation towards sampling plans from this definable space [@fifield2020mcmc; @carter2019; @deford2019; @mccartan2020].

Using sampling methods, we can then develop a reliable counterfactual that can be tuned to consider the sampling constraints imposed by the laws and practices of a state.
This counterfactual allows us to characterize the likely plans from a state.
Often, we can would then call this an ensemble analysis, which is a flexible approach to sampling redistricting plans by simulation, relying on large numbers of plans for stability \cite{matt:etal:21} or sufficient diversity of nearly-independent plans \cite{kenn:etal:forth}.
using this counterfactual, we thus create measures of harm, utility, and fairness which are based on a single plan compared to a set of other reasonable and feasible plans which incorporate the necessary legal constraints to make a plan viable.

We offer a reminder that geographic context is key to understanding common problems in political science..
Elections themselves are geographically determined.
Recent research offers important resolution of the degree of partisan sorting in the US \cite{brow:enos:something}.
By combining those understandings with the necessities of redistricting, we are able to expand the localized aspect of harm.
Further, we suggest that while the efficiency gap may not incorporate geographic information, it often helps capture individual harm better than other common measures.
This is a great benefit as an indicator to detect extreme gerrymanders in cases where sufficiently local geographic data to simulate plans is not available.
We further clarify that common existing measures of partisan fairness -- while they may capture the difficulties offered to a party -- do not necessarily capture the representational harms incurred by individuals who live in districts which do not at all represent their preferences.
